---
title: "Data Analysis CA 4 - Linear Regression Analysis"
author: "James Bunt (D00262403)"
date: "January 2023"
output:
  html_document: null
  pdf_document: default
subtitle: "Part B: Linear Regression Analysis"
---

```{r setup and library load, include=FALSE}

# Clear the console
cat('\014')

# Load necessary packages
library("ggplot2")
```

---

## Introduction
I will be performing simple linear regression from a survey of gamer preferences across 250 undergraduate students. Of the 11 columns within the dataset, 2 columns will be used in this analysis on a random sample of 200 records. I will be analysing whether the two variables 'age' (independent variable) and 'avg_monthly_hrs_gaming' (dependent variable) are correlated.

## Assumptions
Simple linear regression is a parametric test, meaning that it makes certain assumptions about the data. These assumptions are:

1. Homogeneity of variance (homoscedasticity): the size of the error in our prediction doesn’t change significantly across the values of the independent variable, which is validated in this data.
2. Independence of observations: the observations in the dataset were collected using statistically valid sampling methods, and there are no hidden relationships among observations, which is validated in this data.
3. Normality: The data follows a normal distribution.
4. Relationship between the independent and dependent variable is linear.

In summary, this data does meet the above assumptions of homoscedasticity and normality therefore analysis can commence as follows.

---

## Loading and Randomly Sampling the Data
```{r Load Data, echo=TRUE}

# Load data from the csv file
file_name <- "./amalgamated_game_survey_250_2022.csv"
game_survey_data <- read.csv(file_name)

# Set random seed for reproducibility and randomness
set.seed(sample(1:1e+05, 1))

# Select a random sample of 200 records from the data
sample_rows <- sample(1:nrow(game_survey_data), 200, replace = FALSE)
sample <- game_survey_data[sample_rows, ]
```

---

## 1. Histogram of 'age' versus 'average monthly hours gaming'
```{r Histogram, echo=TRUE}
# Create a base plot using the sample data
ggplot(data = sample, aes(x = age)) +
# Add a histogram layer to the plot
geom_histogram(bins = 30, fill = "lightblue") +
# Add labels to the plot
labs(x = "Age", y = "Frequency", title = "Histogram of age")


# Create a base plot using the sample data
ggplot(data = sample, aes(x = avg_monthly_hrs_gaming)) +
# Add a histogram layer to the plot
geom_histogram(bins = 30, fill = "lightblue") +
# Add labels to the plot
labs(x = "Average monthly hours spent gaming", y = "Frequency", title = "Histogram of average monthly hours spent gaming")

```

## 2: Sharpio-Wilk test

The data is uniform as the data points are spread out evenly across the graph.

```{r}

# Load necessary libraries
library(ggplot2)
library(readxl)

# Shapiro-Wilk test for age variable
shapiro.test(sample$age)

# Shapiro-Wilk test for avg_monthly_hrs_gaming variable
shapiro.test(sample$avg_monthly_hrs_gaming)

# create normal probability plot for age variable
qqnorm(sample$age)
qqline(sample$age)

# create normal probability plot for avg_monthly_hrs_gaming variable
qqnorm(sample$avg_monthly_hrs_gaming)
qqline(sample$avg_monthly_hrs_gaming)
```

---


